Community-Conditioned Language Models

### 2020-11-20 paper/paper.tex
checking notes and consistency

We also note that the conditioned models perform differently across different communities---%
for some communities, perplexity of the conditioned models is a lot lower than the unconditioned models,
whereas for others it is the same or even slightly higher (\cref{tab:model-ppls}).\bn{talk about this in terms of mutual information instead?}\jp{What is that?}
As an informal observation, we note 
that across all the models we tested, 
communities where conditioning has the least effect
tend to be more organized around more general interest topics,
such as the \emph{relationships} subreddit, where the subject matter is relevant to a broad
range of people.
Conditioning the model on community has the most benefit
for narrower special-interest subreddits such as those organized
around a certain videogame, sports team, or subculture.
This makes sense intuitively, since communities with more niche topics
would tend to have more specialized vocabulary and other linguistic patterns.

@@

We consider the average information gain of each CCLM over its baseline
un-conditioned counterpart with the same sequence encoder architecture.
For a given message, the information gain is defined as the ratio of the
perplexity of the baseline model to that of the CCLM:
\[
  IG_{CCLM}(m) = \frac{Ppl_{LM}(m)}{Ppl_{CCLM}(m)}
\]

\begin{table*}
  \centering
  % \input{floats/info_gain.tex}
  \caption{Mean information gain of the CCLM over the baseline model, for messages in the test set.}
  \label{tab:model-info-gain}
\end{table*}

@@

As a quantitative assessment, we compare the CCLM-learned community
embedding with the community embedding created by \citet{Kumar2018},
who generated them from the message-community co-occurence matrix,
using data which spans January 2014 to April 2017. We refer the reader
to \citet{Kumar2018}\jp{I actually cannot find in their paper what
  they do.} for details, but the important point is that no linguistic
information is used to create these embeddings: they only depend on
whether and how much a given user participates in a given
community. In sharp contrasts, CCLM community embeddings depend in no
way on which user is the author of any given message: we only use the
contents of messages, not authorship data. 
%
We correlate the pairwise cosine similarity of the
CCLM embeddings with the same metric on the membership embedding,
restricted to the 46 communities chosen for our study
(\cref{fig:pairwise-comm-sim}).

@@

The perplexity of a CCLM gives us a measure of linguistic complexity,
but it is not a measure of how linguistically \emph{specific} the
community is. However, the CCLM induces a latent classifier,
which we call a language model-based community classifier
(LMCC). Using the Bayes theorem, we can turn a
CCLM into a LMCC: the probability that a given message $m$
belongs to a community $c_i$ is calculated as follows:
\[P(c=c_i | m) = P(m | c_i)\frac {P(m)} {P(c=c_i)}\]
In the above,
$P(c=c_i)$ is the frequency of community $c_i$ in the dataset, and
$P(m | c_i)$ (the probability that a message $m$
belongs in a community $c_i$. ) is estimated by the CCLM. $P(m)$ is the absolute
probability of a message, which can be computed as
\[P(m) = \frac 1 {\sum_i P(c=c_i) P(m|c=c_i ) }\].

We can characterise the quality the LMCC with a confusion matrix
$C$, defined by $C_{ij} = P(c=c_i | m)$ for a given (test set) message $m$ 
from community $c_j$. 
That is: \[C_{ij} = average_{m \in Messages(c_j)}(P(c=c_i | m))\].
As a sanity check, we observe that the diagonal dominates the matrix: on average the LMCC picks the right community for a message (\cref{fig:confusion}).

@@

However, we observe that community-wise CCLM perplexity is only weakly (if at all) \bn{Changed the indiscernability (y-axis) from entropy to Ppl and the p-values all shot up above 0.1. However the r is still very weakly positive for every model.}
correlated with community indiscernibility under the corresponding latent 
LMCC (see \cref{fig:cclm_lmcc_ppl} and \cref{tab:cclm_lmcc_ppl}).
In other words, the linguistic complexity of a given community
is a poor predictor of how distinct its language is.

@@

\begin{table}
  \centering
  \input{floats/cclm_lmcc_ppl.tex}
  \caption{
  Pearson's r correlation between community-wise linguistic perplexity and indescernibility for each model.
  }
  \label{tab:cclm_lmcc_ppl}
\end{table}

@@

\begin{table}
  \centering
  % \input{floats/lmcc_gain.tex}
  \caption{
    Pearson's r correlation between community-wise information gain (of the CCLM over the un-conditioned model) and indescernibility.
  }
  \label{tab:lmcc_gain}
\end{table}

@@

With indiscernibility values of 1.5 to 3.5 (out of a maximum of 46),
we find that CCLMs can be used to recognize communities based on
latent bayesian classification \ref{hyp:LMCC-works}. Even the most
generic communities appear to exhibit fairly specific linguistic
patterns. Furthermore, these patterns are just as recognizable by LSTM
or Transformer models, when properly configured and trained.

@@

The representation of communities built by CCLM exhibits a clear
correlation with user-coocurrence patterns
(\ref{hyp:extra-linguistic-correlation}). A natural
question that may arise is how much this correlation can be explain by
user-specifc linguistic patterns. We plan to control for this variable
in future work, for example by building author-conditioned language
models and testing if community vector embeddings can be determined by
the set of the vector embeddings of its members.

>>>>

We also note that the conditioned models perform differently across different communities---%
for some communities, perplexity of the conditioned models is a lot lower than the unconditioned models,
whereas for others it is the same or even slightly higher (\cref{tab:model-ppls}).\bn{talk about this in terms of mutual information instead?}\jp{This should make no difference (but I'm not sure on how you are taking averages; see other comment.)}
As an informal observation, we note 
that across all the models we tested, 
communities where conditioning has the least effect
tend to be more organized around more general interest topics,
such as the \emph{relationships} subreddit, where the subject matter is relevant to a broad
range of people.
Conditioning the model on community has the most benefit
for narrower special-interest subreddits such as those organized
around a certain videogame, sports team, or subculture.
This makes sense intuitively, since communities with more niche topics
would tend to have more specialized vocabulary and other linguistic patterns.

@@

We consider the average information gain of each CCLM over its baseline
un-conditioned counterpart with the same sequence encoder architecture.
For a given message, the information gain is defined as the ratio of the
perplexity of the baseline model to that of the CCLM:
\[
  IG_{CCLM}(m) = \frac{Ppl_{LM}(m)}{Ppl_{CCLM}(m)}
\]
\jp{What we had before: average entropy gain is $H(m)-H_0(m)$, and
  then taking the average over all messages:
  $avg(H(m)-H_0(m)) = avg(H(m))-avg(H_0(m))$. Taking the exponential
  gives a ratio of perplexity.  Here it seems that you have commuted
  exponential and average. It is rather weird to do average in the
  exponential space without normalising to a probability.}
\begin{table*}
  \centering
  % \input{floats/info_gain.tex}
  \caption{Mean information gain of the CCLM over the baseline model, for messages in the test set.}
  \label{tab:model-info-gain}
\end{table*}

@@

As a quantitative assessment, we compare the CCLM-learned community
embedding with the community embedding created by \citet{Kumar2018},
who generated them from the message-community co-occurence matrix,
using data which spans January 2014 to April 2017. We refer the reader
to \citet{Kumar2018} % JP: I actually cannot find in their paper what they do.
for details, but the important point is that no linguistic
information is used to create these embeddings: they only depend on
whether and how much a given user participates in a given
community. In sharp contrasts, CCLM community embeddings depend in no
way on which user is the author of any given message: we only use the
contents of messages, not authorship data. 
%
We correlate the pairwise cosine similarity of the
CCLM embeddings with the same metric on the membership embedding,
restricted to the 46 communities chosen for our study
(\cref{fig:pairwise-comm-sim}).

@@

The perplexity of a CCLM gives us a measure of linguistic complexity,
but it is not a measure of how linguistically \emph{specific} the
community is. However, the CCLM induces a latent classifier,
which we call a language model-based community classifier
(LMCC). Using the Bayes theorem, we can turn a
CCLM into a LMCC: the probability that a given message $m$
belongs to a community $c_i$ is calculated as follows:
\[P(c=c_i | m) = P(m | c_i)\frac {P(m)} {P(c=c_i)}\]
In the above,
$P(c=c_i)$ is the frequency of community $c_i$ in the dataset, and
$P(m | c_i)$ (the probability that a message $m$
belongs in a community $c_i$. ) is estimated by the CCLM. $P(m)$ is the absolute
probability of a message, which can be computed as
\[P(m) = \frac 1 {\sum_i P(c=c_i) P(m|c=c_i ) }\]

We can characterise the quality the LMCC with a confusion matrix
$C$, defined by $C_{ij} = P(c=c_i | m)$ for a given (test set) message $m$ 
from community $c_j$.
\[C_{ij} = average_{m \in Messages(c_j)}(P(c=c_i | m))\]
As a sanity check, we observe that the diagonal dominates the matrix: on average the LMCC picks the right community for a message (\cref{fig:confusion}).

@@

However, we observe that community-wise CCLM perplexity is only weakly (if at all) \bn{Changed the indiscernability (y-axis) from entropy to Ppl and the p-values all shot up above 0.1. However the r is still very weakly positive for every model.}\jp{Are indiscernability values correct? Seems inconsistent with the confusion matrix.}
correlated with community indiscernibility under the corresponding latent 
LMCC (see \cref{fig:cclm_lmcc_ppl} and \cref{tab:cclm_lmcc_ppl}).
In other words, the linguistic complexity of a given community
is a poor predictor of how distinct its language is.

@@

\begin{table}
  \centering
  \input{floats/cclm_lmcc_ppl.tex}
  \caption{
  Pearson's $r$ correlation coefficient between community-wise linguistic perplexity and indescernibility for each model.
  }
  \label{tab:cclm_lmcc_ppl}
\end{table}

@@

\begin{table}
  \centering
  \input{floats/lmcc_gain.tex}
  \caption{
    Pearson's \(r\) correlation coefficient between community-wise information gain (of the CCLM over the un-conditioned model) and indescernibility.
  }
  \label{tab:lmcc_gain}
\end{table}

@@

With indiscernibility values of 1.5 to 3.5 (out of a maximum of 46)\jp{Inconsistent with data; to revisit},
we find that CCLMs can be used to recognize communities based on
latent bayesian classification \ref{hyp:LMCC-works}. Even the most
generic communities appear to exhibit fairly specific linguistic
patterns. Furthermore, these patterns are just as recognizable by LSTM
or Transformer models, when properly configured and trained.

@@

The representation of communities built by CCLM exhibits a clear
correlation with user-coocurrence patterns
\ref{hyp:extra-linguistic-correlation}. A natural
question that may arise is how much this correlation can be explain by
user-specifc linguistic patterns. We plan to control for this variable
in future work, for example by building author-conditioned language
models and testing if community vector embeddings can be determined by
the set of the vector embeddings of its members.
